---
title: "Covid Vaccination Analysis project"
author: "Oviyaa"
date: "31/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Task Details Following are the inspirations to work in this dataset :

How many states have fully vaccinated?
What is the number of positive cases rising in India?
Which states are more affected by the virus?
Analyze to show insights regarding coronavirus

1.Install Packages
```{r}

library(readr)
library(tidyverse)
library(ggplot2)
library(dplyr)
```
2.Read Input file and store it in dataframe
```{r}
############################# Read Covid_19_india data into the datatframe Covid19India #######################################
print("Read Covid_19_india data into the dataframe Covid19India\n")
Covid19India <- read_csv('~/R/data/covid19_data/covid_19_india.csv')
print("How does the data look like")
head(Covid19India)
print("The columns in the data")
colnames(Covid19India)

print("We can drop 2 columns ['ConfirmedIndianNational', 'ConfirmedForeignNational']\n")
Covid19India <- subset(Covid19India, select = -c(ConfirmedIndianNational, ConfirmedForeignNational))
head(Covid19India)

Covid19India['Date'] = as.POSIXct(Covid19India$Date, format = '%Y-%m-%d')
print("No of States: ")
#number of unique states
length(unique(Covid19India$State))

Covid19India<- rename(Covid19India,"State" = "State/UnionTerritory")

Covid19India$State

print("The states are: ")
#unique indian states
unique(Covid19India$State)
```
There are 2 instances of 'State' that we shall drop, 'Cases being reassigned to states' and 'Unassigned'
```{r}

Covid19India<-Covid19India[!grepl('Unassigned',Covid19India$State),]
Covid19India<-Covid19India[!grepl('Cases being reassigned to states',Covid19India$State),]

print("After dropping 'Cases being reassigned to states' and 'Unassigned,the dataframe has :")
str(Covid19India)
```

Sorting Covid19India in descending order of 'Date' and 'Confirmed' into Covid19ToDate
```{r}
Covid19ToDate <- Covid19India[order(Covid19India$Date,Covid19India$Confirmed),decreasing =TRUE]

Covid19ToDate <- subset(Covid19ToDate, select = -c(Sno))
head(Covid19ToDate)
```

Plot the curve of Confirmed, Cured and Death cases in India from Feb 2020 till date
```{r}
Covid19ByDay <- Covid19ToDate %>% group_by(Date) %>% summarise_at(vars(Cured,Deaths,Confirmed),list(sum)) 
Covid19ByDay$CuredPerDay= Covid19ByDay$Cured - lag(Covid19ByDay$Cured,1,default = NA)
Covid19ByDay$DeathsPerDay  = Covid19ByDay$Deaths -lag(Covid19ByDay$Deaths,1,reminder = TRUE)
Covid19ByDay$ConfirmedPerDay  = Covid19ByDay$Confirmed -lag(Covid19ByDay$Confirmed,1,reminder = TRUE)
Covid19ByDay
View(Covid19ByDay)

Covid19ByDay = filter(Covid19ByDay,Covid19ByDay$Date >= '2020-02-01')

##Visualization using ggplot
ggplot(data = Covid19ByDay) + geom_line(mapping=aes(x=Date,y=Confirmed, color = 'dodgerblue'))+geom_line(mapping=aes(x=Date,y=Cured, color = 'green'))+geom_line(mapping=aes(x=Date,y=Deaths, color = 'red'))+ labs(y ="Cumulative graph In Crores", fontsize=16)

Covid19ByDay1 = filter(Covid19ByDay,Covid19ByDay$Date >= '2021-01-01')

##Visualization using ggplot
ggplot(data = Covid19ByDay1) + geom_line(mapping=aes(x=Date,y=Confirmed, color = 'dodgerblue'))+geom_line(mapping=aes(x=Date,y=Cured, color = 'green'))+geom_line(mapping=aes(x=Date,y=Deaths, color = 'red'))+ labs(y ="Cumulative graph In Crores", fontsize=16)

Latest_date <- max(Covid19ToDate$Date)
CovidWorstStates <- Covid19ToDate %>% filter(Date == Latest_date)
nrow(CovidWorstStates)

CovidWorstStates<- CovidWorstStates[order(-CovidWorstStates$Confirmed),]
CovidWorstStates

CovidWorstStates_5 <- head(CovidWorstStates,5)
CovidWorstStates_5

ggplot(data = CovidWorstStates_5,mapping=aes(x=State,y=Confirmed,fill=State)) + geom_bar(stat="identity")+ labs(y ="Bar graph on which state affected most", fontsize=16)+theme(axis.text.x = element_text(angle=90)) 


barplot(CovidWorstStates_5$Confirmed,names.arg=CovidWorstStates_5$State,xlab="States",ylab="Confirmed Cases", col="blue")

barplot(CovidWorstStates_5$Cured,names.arg=CovidWorstStates_5$State,xlab="States",ylab="Cured Cases", col="blue")

barplot(CovidWorstStates_5$Deaths,names.arg=CovidWorstStates_5$State,xlab="States",ylab="Death Cases", col="blue")
```
Read covid_vaccine_statewise.csv
```{r}
covid_vaccine_statewise <- read_csv("~/R/data/covid19_data/covid_vaccine_statewise.csv")
head(covid_vaccine_statewise)
colnames(covid_vaccine_statewise)
str(covid_vaccine_statewise)
CovidVaccine <-covid_vaccine_statewise
```
##Data Preparation
```{r}

#Rename columns

CovidVaccine<-rename(CovidVaccine,'Date'='Updated On',
                                          'TotalVaccinated'=  'Total Individuals Vaccinated' ,
                                           'FirstDose'='First Dose Administered',
                                           'SecondDose' = 'Second Dose Administered',
                                           'Male'='Male(Individuals Vaccinated)',
                                           'Female'='Female(Individuals Vaccinated)',
                                           'Transgender'='Transgender(Individuals Vaccinated)',
                                          'TotalCovaxin'= 'Covaxin (Doses Administered)',
                                           'TotalCoviShield'='CoviShield (Doses Administered)')

#no of not null values
colSums(is.na(CovidVaccine))

################################# We drop the rows with Nan values in 'TotalVaccinated' ######################################
CovidVaccine <- CovidVaccine %>% drop_na(TotalVaccinated)

################################## Convert datetime and get the latest date #################################################
CovidVaccine$Date = as.POSIXct(CovidVaccine$Date, format = "%d/%m/%Y") 
head(CovidVaccine)

latest_date = max(CovidVaccine$Date)

print("Latest Date : ") 
latest_date
```

```{r}
unique(CovidVaccine$State)
#number of unique states
length(unique(CovidVaccine$State))


filter(CovidVaccine,State == "Dadra and Nagar Haveli and Daman and Diu")

filter(CovidVaccine,State == "Jammu and Kashmir + Ladakh")

############################################################################################################################
##################### We change the name of CovidVaccine.State to match with StatePop.State ################################

CovidVaccine$State[920:1031] = 'Dadra and Nagar Haveli & Daman and Diu'
CovidVaccine$State[1610:1723] = 'Jammu and Kashmir + Ladakh'
CovidVaccine$State[2990:3103] = 'Orissa'
```

```{r}
print("After some data cleaning and organising, the rows with Nan values")
colSums(is.na(CovidVaccine))
```
##Covid Vaccine Viualization
```{r}
CovidVaccine_plot<-CovidVaccine %>% filter(CovidVaccine$State=="India")

ggplot(data=CovidVaccine_plot)+geom_line(mapping = aes(x=Date, y=TotalVaccinated))
```
##How the Better Vaccinated States are doing in the Corona 
```{r}
CovidVaccine_state_group <- CovidVaccine %>% filter(Date==latest_date)
CovidVaccine_state_group 
```

Delete Row with India store it in a variable
```{r}
CovidVaccine_India <- CovidVaccine %>% filter(State=="India",Date==latest_date)
CovidVaccine_India

##Delete India as row in CovidVaccine_state_group 
CovidVaccine_state_group <- CovidVaccine_state_group[!(CovidVaccine_state_group$State=="India"),]

#List State names and Total vaccinated count
CovidVaccine_state_Vacc_count<-CovidVaccine_state_group %>% select(State,TotalVaccinated)
CovidVaccine_state_Vacc_count

#Arrange the states in descending order
CovidVaccine_state_Vacc_count<- CovidVaccine_state_Vacc_count[order(-CovidVaccine_state_Vacc_count$TotalVaccinated),]
CovidVaccine_state_Vacc_count

#Top 10 Vaccinates States in India
top_10_vacc=head(CovidVaccine_state_Vacc_count,10)
top_10_vacc
```
##Viz of top 10 vaccinated States
```{r}

ggplot(data=top_10_vacc,mapping = aes(x=State,y=TotalVaccinated,fill=TotalVaccinated))+ geom_bar(stat="identity")+ theme(axis.text.x = element_text(angle=90), plot.title = element_text(hjust = 0.5)) + labs(title = "Top 10 vaccinated States",subtitle = "Analysis done as per latest_date",caption=latest_date)+geom_text(aes(x = State, y = TotalVaccinated, label = TotalVaccinated)) + coord_flip() 
```
## How the Better Vaccinated States are doing in the Corona 
```{r}
CovidStates_cases<- CovidWorstStates[order(-CovidWorstStates$Confirmed),]
#CovidStates_cases-> has list of states with confimed, cured, death number
#CovidVaccine_state_group -> has list of last states with total vaccinated count
CovidStates_cases
CovidVaccine_state_group

#merge two datasets based on state name
CovidVaccine_merge<-merge(CovidVaccine_state_group, CovidStates_cases, by.x="State", by.y="State")
CovidVaccine_merge<-CovidVaccine_merge[order(CovidVaccine_merge$TotalVaccinated),]

#Viz TotalVaccinated vs Confirmed
print("How the Better Vaccinated States are doing in Corona?\n")

ggplot(head(CovidVaccine_merge,10)) + 
  geom_bar(mapping = aes(x = State, y = TotalVaccinated,fill=TotalVaccinated), stat = "identity", position = "dodge") + 
  geom_bar(mapping = aes(x = State, y = Confirmed, fill=Confirmed), stat = "identity", position = "dodge") + coord_flip()+labs(title="Top 10 vaccinated states vs Covid Confirmed cases")

ggplot(tail(CovidVaccine_merge,5)) + 
  geom_bar(mapping = aes(x = State, y = TotalVaccinated,fill=TotalVaccinated), stat = "identity", position = "dodge") + 
  geom_bar(mapping = aes(x = State, y = Confirmed, fill=Confirmed), stat = "identity", position = "dodge") + coord_flip()+labs(title="Last 10 vaccinated states vs Covid Confirmed cases")
```


